---
title: "Structural inequalities in General Practice in England"
format:
  html:
    code-fold: true
    code-summary: "Click to show code"
    fig-width: 90
    fig-height: 25
execute: 
  echo: true
---

```{r}
#| label: setup
#| warning: false
#| echo: true
#| code-fold: true
#| code-summary: "Show the set up code"

# Load necessary libraries
library(ggplot2)
library(tidyverse)
library(magrittr)

# Set up the ggplot theme
theme_set(
  theme_minimal() +
    theme(
      axis.title = element_text(size = 80),
      axis.text = element_text(size = 80),
      plot.caption = element_text(size = 60),
      plot.title = element_text(size = 100),
      plot.subtitle = element_text(size = 90),
      panel.grid = element_line(size = 5),
      legend.title = element_text(size = 60),
      legend.text = element_text(size = 60),
      legend.key.width = unit(3.5, "cm"),
      plot.title.position = "plot"
    )
)

update_geom_defaults("point", list(size = 50))

df <- read.csv("https://raw.githubusercontent.com/HealthEquityEvidenceCentre/HEEC/refs/heads/main/datapacks/final_data.csv") |> 
  mutate(quin_5 = ifelse(is.na(quin_5), quin_4, quin_5))

```

## Introduction

-   Strong primary care is associated with more equitable health outcomes.

-   A key role of commissioners is to ensure the equitable distribution of resources across the system.

-   We present the latest NHS primary care data, using [Index of Multiple Deprivation (IMD)](#imd) to examine inequalities existing in primary care access, experience and outcomes, across the following categories:

    -   **Resources (supply)**: Payments, Workforce
    -   **Population (demand)**: Disease prevalence, Health-related behaviours
    -   **Service quality**: QOF achievement
    -   **Access**: Patient experience, Appointments
    -   **Impact on secondary care**: Emergency admissions, A&E attendances

-   This analysis was produced by the [Health Equity Evidence Centre](https://www.heec.co.uk/). Additional data and analysis is available on [GitHub](https://github.com/HealthEquityEvidenceCentre/HEEC).

-   For further information or to discuss the results, please contact [Dr John Ford](mailto:j.a.ford@qmul.ac.uk) or [Mr Cameron Appel](mailto:c.appel@qmul.ac.uk).

## Inequality in Life Expectancy

```{r Life_Expectancy}
year <- 2019

df_filter <- df %>%
  filter(ICB.NAME == "England") %>%
  filter(Year == year) %>%
  filter(Indicator %in% c("Life_Expectancy_Female", "Life_Expectancy_Male"))

df_filter %>%
  ggplot(., aes(x = Indicator)) +
  geom_linerange(aes(x = Indicator, ymin = quin_5, ymax = quin_1)) +
  geom_point(aes(y = quin_1, color = "Least deprived")) +
  geom_point(aes(y = quin_5, color = "Most deprived")) +
  geom_point(aes(y = avg, color = "National average"), size = 15, shape = 18) +
  scale_x_discrete(labels = c(
    "Life_Expectancy_Female" = "Female",
    "Life_Expectancy_Male" = "Male"
  )) +
  scale_color_manual(values = c("Least deprived" = "#009639", "Most deprived" = "#003087", "National average" = "#DA291C")) +
  labs(
    x = NULL, y = NULL,
    title = paste0("Life expectancy by practice, birth cohort 2016-20 (England)"),
    subtitle = "Practices in most and least deprived IMD quintiles. Larger circles represent stronger socio-economic gradient.",
    colour = "",
    caption = "Source: NHS England GP Workforce statistics, 2024. © Health Equity Evidence Centre, 2024."
  ) +
  coord_flip() +
  guides(colour = guide_legend(override.aes = list(size = 50, shape = c(16, 16, 18))))
```

Average life expectancy for males is **`r round(df_filter[df_filter$Indicator == "Life_Expectancy_Male", ]$quin_1, 1)`** in the least deprived 20% and **`r round(df_filter[df_filter$Indicator == "Life_Expectancy_Male", ]$quin_5, 1)`** in the most deprived 20%.

Average life expectancy for females is **`r round(df_filter[df_filter$Indicator == "Life_Expectancy_Female", ]$quin_1, 1)`** in the least deprived 20% and **`r round(df_filter[df_filter$Indicator == "Life_Expectancy_Female", ]$quin_5, 1)`** in the most deprived 20%.

## Inequality in NHS Payments

```{r payments}
year <- 2023

df_filter <- df %>%
  filter(ICB.NAME == "England") %>%
  filter(Year == year) %>%
  filter(Indicator %in% c("payment_per_patient_all", "payment_per_patient_disp", "payment_per_patient_non_disp")) %>%
  mutate(Indicator = factor(Indicator, levels = c("payment_per_patient_disp", "payment_per_patient_non_disp", "payment_per_patient_all")))

df_filter %>%
  ggplot(., aes(x = Indicator)) +
  geom_linerange(aes(x = Indicator, ymin = quin_5, ymax = quin_1)) +
  geom_point(aes(y = quin_1, color = "Least deprived")) +
  geom_point(aes(y = quin_5, color = "Most deprived")) +
  geom_point(aes(y = avg, color = "National average"), size = 15, shape = 18) +
  scale_x_discrete(labels = c(
    "payment_per_patient_all" = "All practices",
    "payment_per_patient_disp" = "Dispensing practices",
    "payment_per_patient_non_disp" = "Non-dispensing practices"
  )) +
  scale_color_manual(values = c("Least deprived" = "#009639", "Most deprived" = "#003087", "National average" = "#DA291C")) +
  labs(
    x = NULL, y = NULL,
    title = paste0("Mean payment per weighted patient, 2022/23 (England)"),
    subtitle = "Practices in most and least deprived IMD quintiles",
    colour = "",
    caption = "Source: NHS England Payments to General Practice, 2022/23. © Health Equity Evidence Centre, 2024."
  ) +
  coord_flip() +
  guides(colour = guide_legend(override.aes = list(size = 50, shape = c(16, 16, 18))))
```

Average payment per weighted patient is **£`r round(df_filter[df_filter$Indicator == "payment_per_patient_all", ]$quin_5, 2)`** in the most deprived 20% of practices, versus **£`r round(df_filter[df_filter$Indicator == "payment_per_patient_all", ]$quin_1, 2)`** in the least deprived 20%.

If there are no dispensing practices in the most deprived quintile, we use the second most deprived. If both are missing, we omit the data point from the figure.

## Inequality in Workforce

```{r workforce}
year <- 2024

df_filter <- df %>%
  filter(ICB.NAME == "England") %>%
  filter(Year == year) %>%
  filter(Indicator %in% c("TOTAL_GP_EXTGL_FTE", "TOTAL_LOCUUM_TRN_FTE", "TOTAL_NURSES_FTE", "TOTAL_ADMIN_FTE", "TOTAL_DPC_FTE", "PCN_staff")) %>%
  mutate(Indicator = factor(Indicator, levels = c("PCN_staff", "TOTAL_ADMIN_FTE", "TOTAL_DPC_FTE", "TOTAL_NURSES_FTE", "TOTAL_LOCUUM_TRN_FTE", "TOTAL_GP_EXTGL_FTE")))


df_filter %>%
  ggplot(., aes(x = Indicator)) +
  geom_linerange(aes(x = Indicator, ymin = quin_5, ymax = quin_1)) +
  geom_point(aes(y = quin_1, color = "Least deprived")) +
  geom_point(aes(y = quin_5, color = "Most deprived")) +
  geom_point(aes(y = avg, color = "National average"), size = 15, shape = 18) +
  scale_x_discrete(labels = c(
    "TOTAL_GP_EXTGL_FTE" = "Fully-qualified permanent GPs",
    "TOTAL_LOCUUM_TRN_FTE" = "GPs in training grade & locuums",
    "TOTAL_NURSES_FTE" = "Nurses",
    "TOTAL_ADMIN_FTE" = "Admin/Non-clinical",
    "TOTAL_DPC_FTE" = "Direct patient care",
    "PCN_staff" = "PCN-funded staff (all types)"
  )) +
  scale_color_manual(values = c("Least deprived" = "#009639", "Most deprived" = "#003087", "National average" = "#DA291C")) +
  labs(
    x = NULL, y = NULL,
    title = paste0("Average staff FTE per 10,000 weighted patients, 2023/24 (England)"),
    subtitle = "Practices in most and least deprived IMD quintiles",
    colour = "",
    caption = "Source: NHS England GP Workforce & PCN Workforce, 2023-24. © Health Equity Evidence Centre, 2024."
  ) +
  coord_flip() +
  guides(colour = guide_legend(override.aes = list(size = 50, shape = c(16, 16, 18))))
```

Average fully-qualified GPs FTE per 10,000 weighted patients is **`r round(df_filter[df_filter$Indicator == "TOTAL_GP_EXTGL_FTE", ]$quin_5, 1)`** per weighted patient in the most deprived 20% of practices in England versus **`r round(df_filter[df_filter$Indicator == "TOTAL_GP_EXTGL_FTE", ]$quin_1, 1)`** in the least deprived 20%.

## Inequality in Health-related Behaviours

```{r behaviours, echo = F, warning = F}
year <- 2024

df_filter <- df %>%
  filter(ICB.NAME == "England") %>%
  filter(Year == year) %>%
  filter(Indicator %in% c("Hypertension: QOF prevalence (all ages)", "Obesity: QOF prevalence (18+ yrs)", "Smoking prevalence in adults (15+) - current smokers (QOF)"))

# divide by 100 to get percentage
df_filter$quin_1 <- df_filter$quin_1 / 100
df_filter$quin_5 <- df_filter$quin_5 / 100
df_filter$avg <- df_filter$avg / 100

df_filter %>%
  ggplot(., aes(x = Indicator)) +
  geom_linerange(aes(x = Indicator, ymin = quin_5, ymax = quin_1)) +
  geom_point(aes(y = quin_1, color = "Least deprived")) +
  geom_point(aes(y = quin_5, color = "Most deprived")) +
  geom_point(aes(y = avg, color = "National average"), size = 15, shape = 18) +
  scale_x_discrete(labels = c(
    "Smoking prevalence in adults (15+) - current smokers (QOF)" = "Current smokers (15+ yrs)",
    "Obesity: QOF prevalence (18+ yrs)" = "Obesity (18+ yrs)",
    "Hypertension: QOF prevalence (all ages)" = "Hypertension (all ages)"
  )) +
  scale_y_continuous(
    labels = scales::percent,
    limits = c(0, 0.25)
  ) +
  scale_color_manual(values = c("Least deprived" = "#009639", "Most deprived" = "#003087", "National average" = "#DA291C")) +
  labs(
    x = NULL, y = NULL,
    title = paste0("Average prevalence of health-related behaviours, ", year, " (England)"),
    subtitle = "Practices in most and least deprived IMD quintiles",
    colour = "",
    caption = "Source: NHS England Quality and Outcomes Framework, 2023/24. © Health Equity Evidence Centre, 2024."
  ) +
  coord_flip() +
  guides(colour = guide_legend(override.aes = list(size = 50, shape = c(16, 16, 18))))
```

Average prevalence of current smokers (15+ years) is **`r round(df_filter[df_filter$Indicator == "Smoking prevalence in adults (15+) - current smokers (QOF)", ]$quin_5*100, 1)`%** in the most deprived 20% of practices in England versus **`r round(df_filter[df_filter$Indicator == "Smoking prevalence in adults (15+) - current smokers (QOF)", ]$quin_1*100, 1)`%** in the least deprived 20%.

## Inequality in Disease Prevalence

```{r prevalence, echo = F, warning = F}
year <- 2023

df_filter <- df %>%
  filter(ICB.NAME == "England") %>%
  filter(Year == year) %>%
  filter(Indicator %in% c(
    "Atrial fibrillation: QOF prevalence (all ages)",
    "Asthma: QOF prevalence (6+ yrs)",
    "CHD: QOF prevalence (all ages)", "CKD: QOF prevalence (18+ yrs)", "COPD: QOF prevalence (all ages)", "Dementia: QOF prevalence (all ages)", "Depression: QOF prevalence (18+ yrs)", "Diabetes: QOF prevalence (17+ yrs)", "Epilepsy: QOF prevalence (18+ yrs)", "Heart failure with LVSD: QOF prevalence (all ages)", "Learning disability: QOF prevalence (all ages)", "Mental Health: QOF prevalence (all ages)", "Stroke: QOF prevalence (all ages)"
  ))

# divide by 100 to get percentage
df_filter$quin_1 <- df_filter$quin_1 / 100
df_filter$quin_5 <- df_filter$quin_5 / 100
df_filter$avg <- df_filter$avg / 100

df_filter %>%
  ggplot(., aes(x = Indicator)) +
  geom_linerange(aes(x = Indicator, ymin = quin_5, ymax = quin_1)) +
  geom_point(aes(y = quin_1, color = "Least deprived")) +
  geom_point(aes(y = quin_5, color = "Most deprived")) +
  geom_point(aes(y = avg, color = "National average"), size = 15, shape = 18) +
  scale_x_discrete(labels = c(
    "Atrial fibrillation: QOF prevalence (all ages)" = "Atrial fibrillation (all ages)",
    "Asthma: QOF prevalence (6+ yrs)" = "Asthma (6+ yrs)",
    "CHD: QOF prevalence (all ages)" = "Coronary heart disease (all ages)",
    "CKD: QOF prevalence (18+ yrs)" = "Chronic kidney disease (18+ yrs)",
    "COPD: QOF prevalence (all ages)" = "Chronic obstructive pulmonary disease (all ages)",
    "Depression: QOF prevalence (18+ yrs)" = "Depression (18+ yrs)",
    "Diabetes: QOF prevalence (17+ yrs)" = "Diabetes (17+ yrs)",
    "Epilepsy: QOF prevalence (18+ yrs)" = "Epilepsy (18+ yrs)",
    "Heart failure with LVSD: QOF prevalence (all ages)" = "Heart failure (all ages)",
    "Learning disability: QOF prevalence (all ages)",
    "Mental Health: QOF prevalence (all ages)" = "Severe mental illness (all ages)",
    "Stroke: QOF prevalence (all ages)" = "Stroke (all ages)"
  )) +
  scale_y_continuous(
    labels = scales::percent
  ) +
  scale_color_manual(values = c("Least deprived" = "#009639", "Most deprived" = "#003087", "National average" = "#DA291C")) +
  labs(
    x = NULL, y = NULL,
    title = paste0("Average disease prevalence, ", year, " (England)"),
    subtitle = "Practices in most and least deprived IMD quintiles",
    colour = "",
    caption = "Source: Office for Health Improvements and Disparities National GP Profiles, 2022/23. © Health Equity Evidence Centre, 2024."
  ) +
  theme(
    plot.title = element_text(
      size = 105,
      family = "Poppins"
    ),
    plot.subtitle = element_text(
      size = 90,
      family = "Poppins"
    )
  ) +
  coord_flip() +
  guides(colour = guide_legend(override.aes = list(size = 50, shape = c(16, 16, 18))))
```

Average prevalence of diabetes (17+ years) is **`r round(df_filter[df_filter$Indicator == "Diabetes: QOF prevalence (17+ yrs)", ]$quin_5*100, 1)`%** in the most deprived 20% of practices in England**, versus** **`r round(df_filter[df_filter$Indicator == "Diabetes: QOF prevalence (17+ yrs)", ]$quin_1*100, 1)`%** in the least deprived 20%.

Average prevalence of depression (18+ years) is **`r round(df_filter[df_filter$Indicator == "Depression: QOF prevalence (18+ yrs)", ]$quin_5*100, 1)`%** in the most deprived 20% of practices in England, versus **`r round(df_filter[df_filter$Indicator == "Depression: QOF prevalence (18+ yrs)", ]$quin_1*100, 1)`** **%** in the least deprived 20%.

## Inequality in Quality of Service

```{r quality, echo = F, warning = F}
df_filter <- df %>%
  filter(ICB.NAME == "England") %>%
  filter(Year == 2023 & Indicator %in% c(
    "Last BP reading of patients (<80 yrs, with hypertension), in the last 12 months is <= 140/90 mmHg (denominator incl. PCAs)",
    "IFCC-HbA1c <= 58 mmol/mol in patients with diabetes without frailty (denominator incl. PCAs)"
  ) |
    Year == 2024 & Indicator %in% c(
      "% QOF points achieved",
      "child_imms",
      "Women (25-49 yrs), with a record of cervical screening in the last 3.5 yrs (denominator incl. PCAs)",
      "Women (50-64 yrs), with a record of cervical screening in the last 5.5 yrs (denominator incl. PCAs)"
    )) %>%
  mutate(Indicator = factor(Indicator, levels = c("Women (25-49 yrs), with a record of cervical screening in the last 3.5 yrs (denominator incl. PCAs)", "Women (50-64 yrs), with a record of cervical screening in the last 5.5 yrs (denominator incl. PCAs)", "Last BP reading of patients (<80 yrs, with hypertension), in the last 12 months is <= 140/90 mmHg (denominator incl. PCAs)", "IFCC-HbA1c <= 58 mmol/mol in patients with diabetes without frailty (denominator incl. PCAs)", "child_imms", "% QOF points achieved")))

# divide by 100 to get percentage
df_filter$quin_1 <- df_filter$quin_1 / 100
df_filter$quin_5 <- df_filter$quin_5 / 100
df_filter$avg <- df_filter$avg / 100

df_filter %>%
  ggplot(., aes(x = Indicator)) +
  geom_linerange(aes(x = Indicator, ymin = quin_5, ymax = quin_1)) +
  geom_point(aes(y = quin_1, color = "Least deprived")) +
  geom_point(aes(y = quin_5, color = "Most deprived")) +
  geom_point(aes(y = avg, color = "National average"), size = 15, shape = 18) +
  scale_x_discrete(labels = c(
    "% QOF points achieved" = "QOF points achieved",
    "Women (25-49 yrs), with a record of cervical screening in the last 3.5 yrs (denominator incl. PCAs)" = "Women (25-49 yrs) receiving cervical cancer screen last 3.5yrs",
    "Women (50-64 yrs), with a record of cervical screening in the last 5.5 yrs (denominator incl. PCAs)" = "Women (50-64 yrs) receiving cervical cancer screen last 5.5yrs",
    "child_imms" = "% Children 5y received DTaP/IPV and 2 MMR",
    "Last BP reading of patients (<80 yrs, with hypertension), in the last 12 months is <= 140/90 mmHg (denominator incl. PCAs)" = "Last BP reading of hypertensive patients (<80 yrs) <= 140/90 mmHg",
    "IFCC-HbA1c <= 58 mmol/mol in patients with diabetes without frailty (denominator incl. PCAs)" = "Last HbA1c of diabetic patients < = 68 mmol/mol"
  )) +
  scale_y_continuous(
    labels = scales::percent
  ) +
  scale_color_manual(values = c("Least deprived" = "#009639", "Most deprived" = "#003087", "National average" = "#DA291C")) +
  labs(
    x = NULL, y = NULL,
    title = paste0("Average % achievement of QOF domains, 2023/24", " (England)"),
    subtitle = "Practices in most and least deprived IMD quintiles",
    colour = "",
    caption = "Source: NHS England Quality and Outcomes Framework, 2022/23-2023/24. © Health Equity Evidence Centre, 2024."
  ) +
  coord_flip() +
  guides(colour = guide_legend(override.aes = list(size = 50, shape = c(16, 16, 18))))
```

Average QOF points achieved is **`r round(df_filter[df_filter$Indicator == "% QOF points achieved", ]$quin_5*100, 1)`%** in the most deprived 20% of practices in England, versus **`r round(df_filter[df_filter$Indicator == "% QOF points achieved", ]$quin_1*100, 1)`%** in the least deprived 20%.

## Inequality in Patient Experience

```{r exp, echo = F, warning = F}
df_filter <- df %>%
  filter(ICB.NAME == "England") %>%
  filter(
    Year == 2024 & Indicator %in% c(
      "access_pct",
      "continuity_pct",
      "overall_pct",
      "trust_pct"
    ) |
      Year == 2023 & Indicator %in% c(
        "cqc_rating"
      )
  ) %>%
  mutate(Indicator = factor(Indicator, levels = c("access_pct", "continuity_pct", "trust_pct", "cqc_rating", "overall_pct")))

df_filter %>%
  ggplot(., aes(x = Indicator)) +
  geom_linerange(aes(x = Indicator, ymin = quin_5, ymax = quin_1)) +
  geom_point(aes(y = quin_1, color = "Least deprived")) +
  geom_point(aes(y = quin_5, color = "Most deprived")) +
  geom_point(aes(y = avg, color = "National average"), size = 15, shape = 18) +
  scale_x_discrete(labels = c(
    "overall_pct" = "Overall satisfaction",
    "access_pct" = "Experience of contacting the surgery",
    "continuity_pct" = "Continuity of care",
    "trust_pct" = "Confidence and trust",
    "cqc_rating" = "'Good' or 'Outstanding' CQC rating"
  )) +
  scale_y_continuous(
    labels = scales::percent
  ) +
  scale_color_manual(values = c("Least deprived" = "#009639", "Most deprived" = "#003087", "National average" = "#DA291C")) +
  labs(
    x = NULL, y = NULL,
    title = paste0("Average patient experience, 2023/24 (England)"),
    subtitle = "Practices in most and least deprived IMD quintiles",
    colour = "",
    caption = "Source: Ipsos GP Patient Survey, 2024 & CQC, 2023. © Health Equity Evidence Centre, 2024."
  ) +
  coord_flip() +
  guides(colour = guide_legend(override.aes = list(size = 50, shape = c(16, 16, 18))))
```

Average % of practices receiving 'Good' or 'Outstanding' CQC ratings is **`r round(df_filter[df_filter$Indicator == "cqc_rating", ]$quin_5*100, 1)`**% in the most deprived 20% of practices in England, versus **`r round(df_filter[df_filter$Indicator == "cqc_rating", ]$quin_1*100, 1)`%** in the least deprived 20%.

Average % of patients describing their experience as 'Good' is **`r round(df_filter[df_filter$Indicator == "overall_pct", ]$quin_5*100, 1)`%** in the most deprived 20% of practices in England, versus **`r round(df_filter[df_filter$Indicator == "overall_pct", ]$quin_1*100, 1)`%** in the least deprived 20%.

## Inequality in Appointments

```{r appts, echo = F, warning = F}
year <- 2024

df_filter <- df %>%
  filter(ICB.NAME == "England") %>%
  filter(Year == year) %>%
  filter(Indicator %in% c(
    "Telephone", "Face-to-Face", "DNA"
  ))

df_filter %>%
  ggplot(., aes(x = Indicator)) +
  geom_linerange(aes(x = Indicator, ymin = quin_5, ymax = quin_1)) +
  geom_point(aes(y = quin_1, color = "Least deprived")) +
  geom_point(aes(y = quin_5, color = "Most deprived")) +
  geom_point(aes(y = avg, color = "National average"), size = 15, shape = 18) +
  scale_color_manual(values = c("Least deprived" = "#009639", "Most deprived" = "#003087", "National average" = "#DA291C")) +
  labs(
    x = NULL, y = NULL,
    title = paste0("Median appointments and DNAs per 10,000 weighted patients, March 2024 (England)"),
    subtitle = "Practices in most and least deprived IMD quintiles",
    colour = "",
    caption = "Source: NHS England Appointmets in General Practice, 2024. © Health Equity Evidence Centre, 2024."
  ) +
  theme(
    plot.title = element_text(
      size = 105,
      family = "Poppins"
    ),
    plot.subtitle = element_text(
      size = 95,
      family = "Poppins"
    )
  ) +
  coord_flip() +
  guides(colour = guide_legend(override.aes = list(size = 50, shape = c(16, 16, 18))))
```

Average number of Face-to-Face appointments per 10,000 weighted patients is **`r round(df_filter[df_filter$Indicator == "Face-to-Face", ]$quin_5, 1)`** in the most deprived 20% of practices in England, versus **`r round(df_filter[df_filter$Indicator == "Face-to-Face", ]$quin_1, 1)`** in the least deprived 20%.

## Inequality in Impact on Secondary Care

```{r secondary, echo = F, warning = F}
df_filter <- df %>%
  filter(ICB.NAME == "England") %>%
  filter(Year == 2021 & Indicator %in% c(
    "Emergency admissions (0 to 4 years) - CCG"
  ) |
    Year == 2022 & Indicator %in% c(
      "A&E attendances (0 to 4 years)"
    ) |
    Year == 2023 & Indicator %in% c(
      "Number of emergency admissions with cancer"
    ))

df_filter %>%
  ggplot(., aes(x = Indicator)) +
  geom_linerange(aes(x = Indicator, ymin = quin_5, ymax = quin_1)) +
  geom_point(aes(y = quin_1, color = "Least deprived")) +
  geom_point(aes(y = quin_5, color = "Most deprived")) +
  geom_point(aes(y = avg, color = "National average"), size = 15, shape = 18) +
  scale_x_discrete(labels = c(
    "Number of emergency admissions with cancer" = "Number of emergency admissions with cancer",
    "Emergency admissions (0 to 4 years) - CCG" = "Emergency admissions (0-4 yrs)",
    "A&E attendances (0 to 4 years)" = "A&E attendances (0-4 yrs)"
  )) +
  scale_color_manual(values = c("Least deprived" = "#009639", "Most deprived" = "#003087", "National average" = "#DA291C")) +
  labs(
    x = NULL, y = NULL,
    title = paste0("Median emergency admissions or A&E attendances, latest period (England)"),
    subtitle = "Practices in most and least deprived IMD quintiles",
    colour = "",
    caption = "Source: Office for Health Improvements and Disparities National GP Profiles, 2022/23. © Health Equity Evidence Centre, 2024."
  ) +
  theme(
    plot.title = element_text(
      size = 105,
      family = "Poppins"
    ),
    plot.subtitle = element_text(
      size = 95,
      family = "Poppins"
    )
  ) +
  coord_flip() +
  guides(colour = guide_legend(override.aes = list(size = 50, shape = c(16, 16, 18))))
```

Average number of emergency admissions (0-4 years) is **`r round(df_filter[df_filter$Indicator == "Emergency admissions (0 to 4 years) - CCG", ]$quin_5, 1)`** in the most deprived 20% of practices in England, versus **`r round(df_filter[df_filter$Indicator == "Emergency admissions (0 to 4 years) - CCG", ]$quin_1, 1)`** in the least deprived 20%.

## Acknowledgements

-   The work of the Health Equity Evidence Centre is made possible through seed funding from NHS East of England team.

-   The views expressed in this publication are those of the Health Equity Evidence Centre and not necessarily those of NHS England.